---
title: cli
output:
  md_document:
    variant: markdown_github
    toc: true
    toc_depth: 2
    includes:
      before_body: header.md
  html_document:
    toc: true
    toc_depth: 3
always_allow_html: yes
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README",
  out.width = "100%",
  cache = TRUE
)
# Turn on ANSI colors
options(
    crayon.enabled = TRUE,
    crayon.colors = 256)
crayon::num_colors(forget = TRUE)
asciicast::init_knitr_engine(
  startup = quote({
    library(cli)
    set.seed(1) }),
  echo = TRUE,
  echo_input = FALSE)
# asciicast output defaults
options(
    asciicast_knitr_svg = TRUE,
    asciicast_at = "end",
    asciicast_typing_speed = 0,
    asciicast_padding = 20,
    asciicast_window = FALSE,
    asciicast_omit_last_line = FALSE,
    asciicast_cursor = FALSE,
    width = 100,
    asciicast_rows = "auto",
    asciicast_cols = 100,
    asciicast_end_wait = 0)
```

# Features

* Build a CLI using semantic elements: headings, lists, alerts, paragraphs.
* Theming via a CSS-like language.
* Terminal colors via the [crayon](https://github.com/r-lib/crayon) package.
* Create CLI elements in subprocesses, using the
  [callr](https://github.com/r-lib/callr) package.
* All CLI text can contain interpreted string literals, via the
  [glue](https://github.com/tidyverse/glue) package.

# Installation

Install the stable version from CRAN:

```r
install.packages("cli")
```

# Building a command line interface

```{r, cache = FALSE}
library(cli)
```

To start building a CLI, so can simply start using the `cli_*` functions
to create various CLI elements. Their exact formatting depends on the
current theme, see 'Theming' below.

## Alerts

Alerts are typically short messages. cli has four types of alerts (success,
info, warning, danger) and also a generic alert type:

```{asciicast}
cli_alert_success("Updated database.")
```

```{asciicast}
cli_alert_info("Reopened database.")
```

```{asciicast}
cli_alert_warning("Cannot reach GitHub, using local database cache.")
```

```{asciicast}
cli_alert_danger("Failed to connect to database.")
```

```{asciicast}
cli_alert("A generic alert")
```

## Text

Text is automatically wrapped to the terminal width.

```{asciicast}
cli_text(cli:::lorem_ipsum())
```

## Paragraphs

Paragraphs break the output. The default theme inserts an empty line
before and after paragraphs, but only a single empty line is added between
two paragraphs.

```{asciicast}
fun <- function() {
  cli_par()
  cli_text("This is some text.")
  cli_text("Some more text.")
  cli_end()
  cli_par()
  cli_text("Already a new paragraph.")
  cli_end()
}
fun()
```

`cli_end()` closes the latest open paragraph (or other open container).

## Auto-closing containers

If a paragraph (or other container, see 'Generic containers' later), is
opened within a function, cli automatically closes it at the end of the
function, by default. So in the previous example the last `cli_end()` call
is not needed. Use `.auto_close = TRUE` in `cli_par()` to leave the
paragraph open after its calling function returns.

## Headings

cli suppports three levels of headings. This is how they look in the default
theme. The default theme adds an empty line before headings, and an empty
line after `cli_h1()` and `cli_h2()`.

```{asciicast}
cli_h1("Heading 1")
```

```{asciicast}
cli_h2("Heading 2")
```

```{asciicast}
cli_h3("Heading 3")
```

## Interpolation

All cli text is treated as a glue template, with special formatters
available (see the 'Inline text formatting' Section):

```{asciicast}
size <- 123143123
dt <- 1.3454
cli_alert_info(c(
  "Downloaded {prettyunits::pretty_bytes(size)} in ",
  "{prettyunits::pretty_sec(dt)}"))
```

## Inline text formatting

To define inline markup, you can use the regular glue braces, and
after the opening brace, supply the name of the markup formatter with a
leading dot, e.g. for emphasised text, you use `.emph`.
Some examples are below, see `?"inline-markup"` for details.

```{asciicast}
fun <- function() {
  cli_ul()
  cli_li("{.emph Emphasized} text")
  cli_li("{.strong Strong} importance")
  cli_li("A piece of code: {.code sum(a) / length(a)}")
  cli_li("A package name: {.pkg cli}")
  cli_li("A function name: {.fn cli_text}")
  cli_li("A keyboard key: press {.kbd ENTER}")
  cli_li("A file name: {.file /usr/bin/env}")
  cli_li("An email address: {.email bugs.bunny@acme.com}")
  cli_li("A URL: {.url https://acme.com}")
  cli_li("An environment variable: {.envvar R_LIBS}")
}
fun()
```

To combine inline markup and string interpolation, you need to add another
set of braces:

```{asciicast}
dlurl <- "https://httpbin.org/status/404"
cli_alert_danger("Failed to download {.url {dlurl}}.")
```

## Lists

cli has three types of list: ordered, unordered and definition lists, see
`cli_ol()`, `cli_ul()` amd `cli_dl()`:

```{asciicast}
cli_ol(c("item 1", "item 2", "item 3"))
```

```{asciicast}
cli_ul(c("item 1", "item 2", "item 3"))
```

```{asciicast}
cli_dl(c("item 1" = "description 1", "item 2" = "description 2",
         "item 3" = "description 3"))
```

Item text is wrapped to the terminal width:

```{asciicast}
cli_ul(c("item 1" = cli:::lorem_paragraph(1, 20),
         "item 2" = cli:::lorem_paragraph(1, 20)))
```

### Adding list items iteratively

Items can be added one by one:

```{asciicast}
fun <- function() {
  lid <- cli_ul()
  cli_li("Item 1")
  cli_li("Item 2")
  cli_li("Item 3")
  cli_end(lid)
}
fun()
```

The `cli_ul()` call creates a list container, and because its items are
not specified, it leaves the container open. Then items can be added one
by one. (The last `cli_end()` is not necessary, because by default
containers auto-close when their calling function exits.)

### Adding text to an item iteratively

`cli_li()` creates a new container for the list item, within the list
container. You can keep adding text to the item, until the
container is closed via `cli_end()` or a new `cli_li()`, which closes
the current item container, and creates another one for the new item:

```{asciicast}
fun <- function() {
  cli_ul()
  cli_li("First item")
  cli_text("This is still the first item")
  cli_li("This is the second item")
}
fun()
```

### Nested lists

To create nested lists, open nested containers:

```{asciicast}
fun <- function() {
  cli_ol()
  cli_li("Item 1")
  ulid <- cli_ul()
  cli_li("Subitem 1")
  cli_li("Subitem 2")
  cli_end(ulid)
  cli_li("Item 2")
  cli_end()
}
fun()
```

In `cli_end(olid)`, the `olid` is necessary, otherwise `cli_end()` would
only close the container of the list item.

## The status bar

cli supports creating a status bar in the last line of the console, if
the terminal supports the carriage return control character to move the
cursor to the beginning of the line. This is supported in all terminals,
in RStudio, Emacs, RGui, R.app, etc. It is not supported if the output is a
file, e.g. typically on CI systems.

`cli_status()` creates a new status bar, `cli_status_update()` updates a
status bar, and `cli_status_clear()` clears it. `cli_status()` returns an
id, that can be used in `cli_status_update()` and `cli_status_clear()` to
refer to the right status bar.

While it is possible to create multiple status bars, on a typical terminal
only one of them can be shown at any time. cli by default shows the one
that was last created or updated.

While the status bar is active, cli can still produce output, as normal.
This output is created "above" the status bar, which is always kept in the
last line of the screen. See the following example:

```{r, include = FALSE, cache = FALSE}
options(asciicast_at = NULL, asciicast_end_wait = 30)
```

```{asciicast}
f <- function() {
  cli_alert_info("About to start downloads.")
  sb <- cli_status("{symbol$arrow_right} Downloading 10 files.")
  for (i in 9:1) {
    Sys.sleep(0.5)
    if (i == 5) cli_alert_success("Already half-way!")
    cli_status_update(id = sb,
      "{symbol$arrow_right} Got {10-i} files, downloading {i}")
  }
  cli_status_clear(id = sb)
  cli_alert_success("Downloads done.")
}
f()
```

```{r, include = FALSE, cache = FALSE}
options(asciicast_at = "end", asciicast_end_wait = 0)
```

# Theming

The looks of the various CLI elements can be changed via themes.
The cli package comes with a simple built-in theme, and new themes can
be added as well.

## Tags, ids and classes

Similarly to HTML document, the elements of a CLI form a tree of nodes.
Each node has exactly one tag, at most one id, and optionally a set of
classes. E.g. `cli_par()` creates a node with a `<p>` tag, `cli_ol()`
creates a node with an `<ol>` tag, etc. Here is an example CLI tree. It
always starts with a `<body>` tag with id `"body"`, this is created
automatically.

```
<body id="body">
  <par>
    <ol>
      <it>
        <span class="pkg">
```

A cli theme is a named list, where the names are selectors based on tag
names, ids and classes, and the elements of the list are style declarations.
For example, the style of `<h1>` tags looks like this in the built-in theme:

```{asciicast}
builtin_theme()$h1
```

See also `?cli::themes` for the reference and `?cli::simple_theme` for
an example theme.

## Generic containers

`cli_div()` is a generic container, that does not produce any output, but
it can add a new theme. This theme is removed when the `<div>` node is
closed. (Like other containers, `cli_div()` auto-closes when the calling
function exits.)

```{asciicast}
fun <- function() {
  cli_div(theme = list (.alert = list(color = "red")))
  cli_alert("This will be red")
  cli_end()
  cli_alert("Back to normal color")
}
fun()
```

## Theming inline markup

The inline markup formatters always use a `<span>` tag, and add the name
of the formatter as a class.

```{asciicast}
fun <- function() {
  cli_div(theme = list(span.emph = list(color = "orange")))
  cli_text("This is very {.emph important}")
  cli_end()
  cli_text("Back to the {.emph previous theme}")
}
fun()
```

# CLI messages

All `cli_*()` functions are implemented using standard R conditions.
For example a `cli_alert()` call emits an R condition with class
`cli_message`. These messages can be caught, muffled, transfered from a
subprocess to the main R process.

When a cli function is called:

1. cli throws a `cli_message` condition.
2. If this condition is caught and muffled (via the `muffleMessage`
   restart), then nothing else happens.
3. Otherwise the `cli.default_handler` option is checked and if this is
   a function, then it is called with the message.
4. If the `cli.default_handler` option is not set, or it is not a function,
   the default cli handler is called, which shows the text, alert, heading,
   etc. on the screen, using the standard R `message()` function.

```{asciicast}
tryCatch(cli_h1("Heading"), cli_message = function(x) x)
suppressMessages(cli_text("Not shown"))
```

# Subprocesses

If `cli_*()` commands are invoked in a subprocess via
`callr::r_session` (see https://callr.r-lib.org), then they are
automatically copied to the main R process:

```{asciicast}
rs <- callr::r_session$new()
rs$run(function() {
  cli::cli_text("This is subprocess {.emph {Sys.getpid()}} from {.pkg callr}")
  Sys.getpid()
})
invisible(rs$close())
```

# Utility functions

## ANSI colors

cli defines wrapper functions to the crayon package, to create ANSI colored
and styled output in the console. `col_*` functions change the foreground
color, `bg_*` functions change the background color, and `style_*` functions
change the style of the text in some way.

These functions concatenate their arguments using `paste0()`, and add
the `ansi_string` class to their result:

```{asciicast}
cat(col_red("This ", "is ", "red."))
```

Foreground colors:

```{asciicast}
cli_ul(c(
  col_black("black"),
  col_blue("blue"),
  col_cyan("cyan"),
  col_green("green"),
  col_magenta("magenta"),
  col_red("red"),
  col_white("white"),
  col_yellow("yellow"),
  col_grey("grey")
))
```

Note that these might actually look different depending on your terminal
theme. Background colors:

```{asciicast}
cli_ul(c(
  bg_black("black background"),
  bg_blue("blue background"),
  bg_cyan("cyan background"),
  bg_green("green background"),
  bg_magenta("magenta background"),
  bg_red("red background"),
  bg_white("white background"),
  bg_yellow("yellow background")
))
```

Text styles:

```{asciicast}
cli_ul(c(
  style_dim("dim style"),
  style_blurred("blurred style"),
  style_bold("bold style"),
  style_hidden("hidden style"),
  style_inverse("inverse style"),
  style_italic("italic style"),
  style_reset("reset style"),
  style_strikethrough("strikethrough style"),
  style_underline("underline style")
))
```

Not all `style_*` functions are supported by all terminals.

Colors, background colors and styles can be combined:

```{asciicast}
bg_white(style_bold(col_red("TITLE")))
```

`make_ansi_style()` can create custom colors, assuming your terminal
supports them. `combine_ansi_styles()` combines several styles into a
function:

```{asciicast}
col_warn <- combine_ansi_styles(make_ansi_style("pink"), style_bold)
col_warn("This is a warning in pink!")
cat(col_warn("This is a warning in pink!"))
```

## Console capabilities

Query the console width:
```{asciicast}
console_width()
```

Query if the console supports ansi escapes:
```{asciicast}
is_ansi_tty()
```

Hide the cursor, if the console supports it (no-op otherwise):
```{asciicast}
ansi_hide_cursor()
ansi_show_cursor()
```

See also `ansi_with_hidden_cursor()`.

Query if the console supports `\r`:
```{asciicast}
is_dynamic_tty()
```

Query if the console supports UTF-8 output:
```{asciicast}
is_utf8_output()
```

# Non-semantic CLI

While the primary way of using cli is the `cli_*()` functions that generate
semantic CLI elements, you can also use cli functions that create
terminal output directly.

## Unicode characters

The `symbol` variable includes some Unicode characters that are often
useful in CLI messages. They automatically fall back to ASCII symbols
if the platform does not support them. You can use these symbols both
with the semantic `cli_*()` functions and directly.

```{asciicast}
options(asciicast_theme = list(font_family =
  c("Monaco", "'Fira Code'", "Consolas", "Menlo",
    "'Bitstream Vera Sans Mono'", "'Powerline Symbols'", "monospace")))
```

```{asciicast}
cli_text("{symbol$tick} no errors  |  {symbol$cross} 2 warnings")
```

Here is a list of all symbols:

```{asciicast}
list_symbols()
```

```{r include = FALSE}
options(asciicast_theme = NULL)
```

Most symbols were inspired by (and copied from) the awesome
[figures](https://github.com/sindresorhus/figures) JavaScript project.

## Rules

### Simple rule, double rule, bars

```{asciicast}
rule()
```

```{asciicast}
rule(line = 2)
```

```{asciicast}
rule(line = "bar2")
```

```{asciicast}
rule(line = "bar5")
```

```{asciicast}
rule(center = "TITLE", line = "~")
```

### Labels

```{asciicast}
rule(left = "Results")
```

```{asciicast}
rule(center = " * RESULTS * ")
```

### Colors

```{asciicast}
rule(center = col_red(" * RESULTS * "))
```

```{asciicast}
rule(center = " * RESULTS * ", col = "red")
```

```{asciicast}
rule(center = " * RESULTS * ", line_col = "red")
```

```{asciicast}
rule(center = "TITLE", line = "~-", line_col = "blue")
```

```{asciicast}
rule(center = bg_red(" ", symbol$star, "TITLE", symbol$star, " "),
  line = "\u2582", line_col = "orange")
```

## Boxes

```{asciicast}
boxx("Hello there!")
```

### Change border style

```{asciicast}
boxx("Hello there!", border_style = "double")
```

### Multiple lines of text

```{asciicast}
boxx(c("Hello", "there!"), padding = 1)
```

### Padding and margin

```{asciicast}
boxx("Hello there!", padding = 1)
```

```{asciicast}
boxx("Hello there!", padding = c(1, 5, 1, 5))
```

```{asciicast}
boxx("Hello there!", margin = 1)
```

```{asciicast}
boxx("Hello there!", margin = c(1, 5, 1, 5))
```

### Floating

```{asciicast}
boxx("Hello there!", padding = 1, float = "center")
```

```{asciicast}
boxx("Hello there!", padding = 1, float = "right")
```

### Colors

```{asciicast}
boxx(col_cyan("Hello there!"), padding = 1, float = "center")
```

```{asciicast}
boxx("Hello there!", padding = 1, background_col = "brown")
```

```{asciicast}
boxx("Hello there!", padding = 1, background_col = bg_red)
```

```{asciicast}
boxx("Hello there!", padding = 1, border_col = "green")
```

```{asciicast}
boxx("Hello there!", padding = 1, border_col = col_red)
```

### Label alignment

```{asciicast}
boxx(c("Hi", "there", "you!"), padding = 1, align = "left")
```

```{asciicast}
boxx(c("Hi", "there", "you!"), padding = 1, align = "center")
```

```{asciicast}
boxx(c("Hi", "there", "you!"), padding = 1, align = "right")
```

### A very customized box

```{asciicast}
star <- symbol$star
label <- c(paste(star, "Hello", star), "  there!")
boxx(
  col_white(label),
  border_style="round",
  padding = 1,
  float = "center",
  border_col = "tomato3",
  background_col="darkolivegreen"
)
```

## Trees

You can specify the tree with a two column data frame, containing the
node ids/labels, and the list of their children.

```{asciicast}
data <- data.frame(
  stringsAsFactors = FALSE,
  package = c("processx", "backports", "assertthat", "Matrix",
    "magrittr", "rprojroot", "clisymbols", "prettyunits", "withr",
    "desc", "igraph", "R6", "crayon", "debugme", "digest", "irlba",
    "rcmdcheck", "callr", "pkgconfig", "lattice"),
  dependencies = I(list(
    c("assertthat", "crayon", "debugme", "R6"), character(0),
    character(0), "lattice", character(0), "backports", character(0),
    c("magrittr", "assertthat"), character(0),
    c("assertthat", "R6", "crayon", "rprojroot"),
    c("irlba", "magrittr", "Matrix", "pkgconfig"), character(0),
    character(0), "crayon", character(0), "Matrix",
    c("callr", "clisymbols", "crayon", "desc", "digest", "prettyunits",
      "R6", "rprojroot", "withr"),
    c("processx", "R6"), character(0), character(0)
  ))
)
tree(data, root = "rcmdcheck")
```

An optional third column may contain custom labels. These can be colored
as well:

```{asciicast}
data$label <- paste(data$package,
  col_grey(paste0("(", c("2.0.0.1", "1.1.1", "0.2.0", "1.2-11",
    "1.5", "1.2", "1.2.0", "1.0.2", "2.0.0", "1.1.1.9000", "1.1.2",
    "2.2.2", "1.3.4", "1.0.2", "0.6.12", "2.2.1", "1.2.1.9002",
    "1.0.0.9000", "2.0.1", "0.20-35"), ")"))
  )
roots <- ! data$package %in% unlist(data$dependencies)
data$label[roots] <- col_cyan(style_italic(data$label[roots]))
tree(data, root = "rcmdcheck")
```

## Spinners

See `list_spinners()` and `get_spinner()`. From the awesome
[cli-spinners](https://github.com/sindresorhus/cli-spinners#readme)
project.

```{asciicast}
list_spinners()
```

```{asciicast}
get_spinner("dots")
```

```{r, include = FALSE, cache = FALSE}
options(asciicast_at = NULL)
```

```{asciicast}
ansi_with_hidden_cursor(demo_spinners("dots"))
```

```{asciicast}
ansi_with_hidden_cursor(demo_spinners("clock"))
```

```{r, include = FALSE, cache = FALSE}
options(asciicast_at = "end")
```

# License

MIT © RStudio
