---
title: ANSI function benchmarks
author: "Gábor Csárdi"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    toc: true
    toc_depth: 2
editor_options:
  markdown:
    wrap: sentence
---

```{r, setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  cache = TRUE
)
```

```{r comment = "", results = 'asis', echo = FALSE}
fansi::set_knit_hooks(
   knitr::knit_hooks,
   which = "output"
)
```

# Introduction

Often we can use the corresponding base R function as a baseline.
We also compare to the fansi package, where it is possible.

# Data

In cli the typical use case is short string scalars, but we run some benchmarks longer strings and string vectors as well.

```{r}
library(cli)
library(fansi)
options(cli.unicode = TRUE)
options(cli.num_colors = 256)
```

```{r}
ansi <- format_inline(
  "{col_green(symbol$tick)} {.code print(x)} {.emph emphasised}"
)
```

```{r}
plain <- ansi_strip(ansi)
```

```{r}
vec_plain <- rep(plain, 100)
vec_ansi <- rep(ansi, 100)
vec_plain6 <- rep(plain, 6)
vec_ansi6 <- rep(plain, 6)
```

```{r}
txt_plain <- paste(vec_plain, collapse = " ")
txt_ansi <- paste(vec_ansi, collapse = " ")
```

```{r}
uni <- paste(
  "\U0001f477\u200d\u2640\ufe0f",
  "\U0001f477\U0001f3fb",
  "\U0001f477\u200d\u2640\ufe0f",
  "\U0001f477\U0001f3fb",
  "\U0001f477\U0001f3ff\u200d\u2640\ufe0f"
)
vec_uni <- rep(uni, 100)
txt_uni <- paste(vec_uni, collapse = " ")
```

# ANSI functions

## `ansi_align()`

```{r}
bench::mark(
  ansi  = ansi_align(ansi, width = 20),
  plain = ansi_align(plain, width = 20), 
  base  = format(plain, width = 20),
  check = FALSE
)
```

```{r}
bench::mark(
  ansi  = ansi_align(ansi, width = 20, align = "right"),
  plain = ansi_align(plain, width = 20, align = "right"), 
  base  = format(plain, width = 20, justify = "right"),
  check = FALSE
)
```

## `ansi_chartr()`

```{r}
bench::mark(
  ansi  = ansi_chartr("abc", "XYZ", ansi),
  plain = ansi_chartr("abc", "XYZ", plain),
  base  = chartr("abc", "XYZ", plain),
  check = FALSE
)
```

## `ansi_columns()`

```{r}
bench::mark(
  ansi  = ansi_columns(vec_ansi6, width = 120),
  plain = ansi_columns(vec_plain6, width = 120),
  check = FALSE
)
```

## `ansi_has_any()`

```{r}
bench::mark(
  cli_ansi        = ansi_has_any(ansi),
  fansi_ansi      = has_sgr(ansi),
  cli_plain       = ansi_has_any(plain),
  fansi_plain     = has_sgr(plain),
  cli_vec_ansi    = ansi_has_any(vec_ansi),
  fansi_vec_ansi  = has_sgr(vec_ansi),
  cli_vec_plain   = ansi_has_any(vec_plain),
  fansi_vec_plain = has_sgr(vec_plain),
  cli_txt_ansi    = ansi_has_any(txt_ansi),
  fansi_txt_ansi  = has_sgr(txt_ansi),
  cli_txt_plain   = ansi_has_any(txt_plain),
  fansi_txt_plain = has_sgr(vec_plain),
  check = FALSE
)
```

## `ansi_html()`

This is typically used with longer text.

```{r}
bench::mark(
  cli   = ansi_html(txt_ansi),
  fansi = sgr_to_html(txt_ansi, classes = TRUE),
  check = FALSE
)
```

## `ansi_nchar()`

```{r}
bench::mark(
  cli_ansi        = ansi_nchar(ansi),
  fansi_ansi      = nchar_sgr(ansi),
  base_ansi       = nchar(ansi),
  cli_plain       = ansi_nchar(plain),
  fansi_plain     = nchar_sgr(plain),
  base_plain      = nchar(plain),
  cli_vec_ansi    = ansi_nchar(vec_ansi),
  fansi_vec_ansi  = nchar_sgr(vec_ansi),
  base_vec_ansi   = nchar(vec_ansi),
  cli_vec_plain   = ansi_nchar(vec_plain),
  fansi_vec_plain = nchar_sgr(vec_plain),
  base_vec_plain  = nchar(vec_plain),
  cli_txt_ansi    = ansi_nchar(txt_ansi),
  fansi_txt_ansi  = nchar_sgr(txt_ansi),
  base_txt_ansi   = nchar(txt_ansi),
  cli_txt_plain   = ansi_nchar(txt_plain),
  fansi_txt_plain = nchar_sgr(txt_plain),
  base_txt_plain  = nchar(txt_plain),
  check = FALSE
)
```

```{r}
bench::mark(
  cli_ansi        = ansi_nchar(ansi, type = "width"),
  fansi_ansi      = nchar_sgr(ansi, type = "width"),
  base_ansi       = nchar(ansi, "width"),
  cli_plain       = ansi_nchar(plain, type = "width"),
  fansi_plain     = nchar_sgr(plain, type = "width"),
  base_plain      = nchar(plain, "width"),
  cli_vec_ansi    = ansi_nchar(vec_ansi, type = "width"),
  fansi_vec_ansi  = nchar_sgr(vec_ansi, type = "width"),
  base_vec_ansi   = nchar(vec_ansi, "width"),
  cli_vec_plain   = ansi_nchar(vec_plain, type = "width"),
  fansi_vec_plain = nchar_sgr(vec_plain, type = "width"),
  base_vec_plain  = nchar(vec_plain, "width"),
  cli_txt_ansi    = ansi_nchar(txt_ansi, type = "width"),
  fansi_txt_ansi  = nchar_sgr(txt_ansi, type = "width"),
  base_txt_ansi   = nchar(txt_ansi, "width"),
  cli_txt_plain   = ansi_nchar(txt_plain, type = "width"),
  fansi_txt_plain = nchar_sgr(txt_plain, type = "width"),
  base_txt_plain  = nchar(txt_plain, type = "width"),
  check = FALSE
)
```

## `ansi_simplify()`

Nothing to compare here.

```{r}
bench::mark(
  cli_ansi      = ansi_simplify(ansi),
  cli_plain     = ansi_simplify(plain),
  cli_vec_ansi  = ansi_simplify(vec_ansi),
  cli_vec_plain = ansi_simplify(vec_plain),
  cli_txt_ansi  = ansi_simplify(txt_ansi),
  cli_txt_plain = ansi_simplify(txt_plain),
  check = FALSE
)
```

## `ansi_strip()`

```{r}
bench::mark(
  cli_ansi        = ansi_strip(ansi),
  fansi_ansi      = strip_sgr(ansi),
  cli_plain       = ansi_strip(plain),
  fansi_plain     = strip_sgr(plain),
  cli_vec_ansi    = ansi_strip(vec_ansi),
  fansi_vec_ansi  = strip_sgr(vec_ansi),
  cli_vec_plain   = ansi_strip(vec_plain),
  fansi_vec_plain = strip_sgr(vec_plain),
  cli_txt_ansi    = ansi_strip(txt_ansi),
  fansi_txt_ansi  = strip_sgr(txt_ansi),
  cli_txt_plain   = ansi_strip(txt_plain),
  fansi_txt_plain = strip_sgr(txt_plain),
  check = FALSE
)
```

## `ansi_strsplit()`

```{r}
bench::mark(
  cli_ansi        = ansi_strsplit(ansi, "i"),
  fansi_ansi      = strsplit_sgr(ansi, "i"),
  base_ansi       = strsplit(ansi, "i"),
  cli_plain       = ansi_strsplit(plain, "i"),
  fansi_plain     = strsplit_sgr(plain, "i"),
  base_plain      = strsplit(plain, "i"),
  cli_vec_ansi    = ansi_strsplit(vec_ansi, "i"),
  fansi_vec_ansi  = strsplit_sgr(vec_ansi, "i"),
  base_vec_ansi   = strsplit(vec_ansi, "i"),
  cli_vec_plain   = ansi_strsplit(vec_plain, "i"),
  fansi_vec_plain = strsplit_sgr(vec_plain, "i"),
  base_vec_plain  = strsplit(vec_plain, "i"),
  cli_txt_ansi    = ansi_strsplit(txt_ansi, "i"),
  fansi_txt_ansi  = strsplit_sgr(txt_ansi, "i"),
  base_txt_ansi   = strsplit(txt_ansi, "i"),
  cli_txt_plain   = ansi_strsplit(txt_plain, "i"),
  fansi_txt_plain = strsplit_sgr(txt_plain, "i"),
  base_txt_plain  = strsplit(txt_plain, "i"),
  check = FALSE
)
```

## `ansi_strtrim()`

```{r}
bench::mark(
  cli_ansi        = ansi_strtrim(ansi, 10),
  fansi_ansi      = strtrim_sgr(ansi, 10),
  base_ansi       = strtrim(ansi, 10),
  cli_plain       = ansi_strtrim(plain, 10),
  fansi_plain     = strtrim_sgr(plain, 10),
  base_plain      = strtrim(plain, 10),
  cli_vec_ansi    = ansi_strtrim(vec_ansi, 10),
  fansi_vec_ansi  = strtrim_sgr(vec_ansi, 10),
  base_vec_ansi   = strtrim(vec_ansi, 10),
  cli_vec_plain   = ansi_strtrim(vec_plain, 10),
  fansi_vec_plain = strtrim_sgr(vec_plain, 10),
  base_vec_plain  = strtrim(vec_plain, 10),
  cli_txt_ansi    = ansi_strtrim(txt_ansi, 10),
  fansi_txt_ansi  = strtrim_sgr(txt_ansi, 10),
  base_txt_ansi   = strtrim(txt_ansi, 10),
  cli_txt_plain   = ansi_strtrim(txt_plain, 10),
  fansi_txt_plain = strtrim_sgr(txt_plain, 10),
  base_txt_plain  = strtrim(txt_plain, 10),
  check = FALSE
)
```

## `ansi_strwrap()`

This function is most useful for longer text, but it is often called for short text in cli, so it makes sense to benchmark that as well.

```{r}
bench::mark(
  cli_ansi        = ansi_strwrap(ansi, 30),
  fansi_ansi      = strwrap_sgr(ansi, 30),
  base_ansi       = strwrap(ansi, 30),
  cli_plain       = ansi_strwrap(plain, 30),
  fansi_plain     = strwrap_sgr(plain, 30),
  base_plain      = strwrap(plain, 30),
  cli_vec_ansi    = ansi_strwrap(vec_ansi, 30),
  fansi_vec_ansi  = strwrap_sgr(vec_ansi, 30),
  base_vec_ansi   = strwrap(vec_ansi, 30),
  cli_vec_plain   = ansi_strwrap(vec_plain, 30),
  fansi_vec_plain = strwrap_sgr(vec_plain, 30),
  base_vec_plain  = strwrap(vec_plain, 30),
  cli_txt_ansi    = ansi_strwrap(txt_ansi, 30),
  fansi_txt_ansi  = strwrap_sgr(txt_ansi, 30),
  base_txt_ansi   = strwrap(txt_ansi, 30),
  cli_txt_plain   = ansi_strwrap(txt_plain, 30),
  fansi_txt_plain = strwrap_sgr(txt_plain, 30),
  base_txt_plain  = strwrap(txt_plain, 30),
  check = FALSE
)
```

## `ansi_substr()`

```{r}
bench::mark(
  cli_ansi        = ansi_substr(ansi, 2, 10),
  fansi_ansi      = substr_sgr(ansi, 2, 10),
  base_ansi       = substr(ansi, 2, 10),
  cli_plain       = ansi_substr(plain, 2, 10),
  fansi_plain     = substr_sgr(plain, 2, 10),
  base_plain      = substr(plain, 2, 10),
  cli_vec_ansi    = ansi_substr(vec_ansi, 2, 10),
  fansi_vec_ansi  = substr_sgr(vec_ansi, 2, 10),
  base_vec_ansi   = substr(vec_ansi, 2, 10),
  cli_vec_plain   = ansi_substr(vec_plain, 2, 10),
  fansi_vec_plain = substr_sgr(vec_plain, 2, 10),
  base_vec_plain  = substr(vec_plain, 2, 10),
  cli_txt_ansi    = ansi_substr(txt_ansi, 2, 10),
  fansi_txt_ansi  = substr_sgr(txt_ansi, 2, 10),
  base_txt_ansi   = substr(txt_ansi, 2, 10),
  cli_txt_plain   = ansi_substr(txt_plain, 2, 10),
  fansi_txt_plain = substr_sgr(txt_plain, 2, 10),
  base_txt_plain  = substr(txt_plain, 2, 10),
  check = FALSE
)
```

## `ansi_tolower()` , `ansi_toupper()`

```{r}
bench::mark(
  cli_ansi        = ansi_tolower(ansi),
  base_ansi       = tolower(ansi),
  cli_plain       = ansi_tolower(plain),
  base_plain      = tolower(plain),
  cli_vec_ansi    = ansi_tolower(vec_ansi),
  base_vec_ansi   = tolower(vec_ansi),
  cli_vec_plain   = ansi_tolower(vec_plain),
  base_vec_plain  = tolower(vec_plain),
  cli_txt_ansi    = ansi_tolower(txt_ansi),
  base_txt_ansi   = tolower(txt_ansi),
  cli_txt_plain   = ansi_tolower(txt_plain),
  base_txt_plain  = tolower(txt_plain),
  check = FALSE
)
```

## `ansi_trimws()`

```{r}
bench::mark(
  cli_ansi        = ansi_trimws(ansi),
  base_ansi       = trimws(ansi),
  cli_plain       = ansi_trimws(plain),
  base_plain      = trimws(plain),
  cli_vec_ansi    = ansi_trimws(vec_ansi),
  base_vec_ansi   = trimws(vec_ansi),
  cli_vec_plain   = ansi_trimws(vec_plain),
  base_vec_plain  = trimws(vec_plain),
  cli_txt_ansi    = ansi_trimws(txt_ansi),
  base_txt_ansi   = trimws(txt_ansi),
  cli_txt_plain   = ansi_trimws(txt_plain),
  base_txt_plain  = trimws(txt_plain),
  check = FALSE
)
```

# UTF-8 functions

## `utf8_nchar()`

```{r}
bench::mark(
  cli        = utf8_nchar(uni, type = "chars"),
  base       = nchar(uni, "chars"),
  cli_vec    = utf8_nchar(vec_uni, type = "chars"),
  base_vec   = nchar(vec_uni, "chars"),
  cli_txt    = utf8_nchar(txt_uni, type = "chars"),
  base_txt   = nchar(txt_uni, "chars"),
  check = FALSE
)
```

```{r}
bench::mark(
  cli        = utf8_nchar(uni, type = "width"),
  base       = nchar(uni, "width"),
  cli_vec    = utf8_nchar(vec_uni, type = "width"),
  base_vec   = nchar(vec_uni, "width"),
  cli_txt    = utf8_nchar(txt_uni, type = "width"),
  base_txt   = nchar(txt_uni, "width"),
  check = FALSE
)
```

```{r}
bench::mark(
  cli        = utf8_nchar(uni, type = "codepoints"),
  base       = nchar(uni, "chars"),
  cli_vec    = utf8_nchar(vec_uni, type = "codepoints"),
  base_vec   = nchar(vec_uni, "chars"),
  cli_txt    = utf8_nchar(txt_uni, type = "codepoints"),
  base_txt   = nchar(txt_uni, "chars"),
  check = FALSE
)
```

## `utf8_substr()`

```{r}
bench::mark(
  cli        = utf8_substr(uni, 2, 10),
  base       = substr(uni, 2, 10),
  cli_vec    = utf8_substr(vec_uni, 2, 10),
  base_vec   = substr(vec_uni, 2, 10),
  cli_txt    = utf8_substr(txt_uni, 2, 10),
  base_txt   = substr(txt_uni, 2, 10),
  check = FALSE
)
```

# Session info

```{r}
sessioninfo::session_info()
```
