
```{r setup, include = FALSE}
pkgload::load_all()
```

# Components summary

## Text pieces

- [x] plain text
- [x] substitution
- [x] `span`

## Inline

- [x] `span`
- [x] `text`

## Block

- [ ] `blockquote`
- [x] `div`
- [ ] `dl`
- [ ] `h1`, `h2`, `h3`
- [ ] `li`
- [x] ~~`meta`~~ Not needed?
- [ ] `ol`
- [ ] `par`
- [ ] `rule`
- [ ] `status` (is this needed here?)
- [ ] `ul`
- [ ] `verbatim`

## Composite components:

- [ ] `alert`: a `div`
- [ ] `alert_danger`: a `div`
- [ ] `alert_info`: a `div`
- [ ] `alert_success`: a `div`
- [ ] `alert_warning`: a `div`
- [ ] `bullets`: using an `ul`

# Component details

## Plain text (text piece)

Plain text piece. It does not have direct styling, but it may have
inherited styling, as a `style` attribute.

### Styles

- [x] `background-color`
- [x] `color`
- [x] `fmt`
- [x] `font-style`
- [x] `font-weight`
- [x] `text-decoration`

## Substitution (text piece)

A substitution. It has a `code` entry, the original code, a `value`
entry, the R object the `code` evaluated to, and a `style` entry that cannot
be specified directly, but can be inherited.

### Styles

This is the only (sub)component that has vector-formatting and
collapsing styles.

Most other styles apply elementwise

- [x] `after` (elementwise)
- [x] `background-color` (elementwise)
- [x] `before` (elementwise)
- [x] ~~`class-map`~~ we can deprecate this, and have a better way to
      define themes for vector classes directly.
- [x] `collapse` from vector to scalar
- [x] `color` (elementwise)
- [x] `digits` This was implemented via a `transform` style and calling
      `cli_vector()` explicitly.
- [x] `fmt` (vectorized)
- [x] `font-style` (elementwise)
- [x] `font-weight` (elementwise)
- [x] `prefix` (elementwise)
- [x] `postfix` (elementwise)
- [x] `string-quote` This was implemented via a `transform` style and calling
      `cli_vector()` explicitly.
- [x] `text-decoration` (elementwise)
- [x] `transform` (vectorized)
- [x] `vec-last`
- [x] `vec-sep`
- [x] `vec-sep2`
- [x] `vec-trunc`
- [x] `vec-trunc-style`

## `span` (inline and text piece)

### Styles

- [x] `after`
- [x] `background-color`
- [x] `before`
- [x] `color`
- [x] `fmt`
- [x] `font-style`
- [x] `font-weight`
- [ ] `margin-left` (need collapsing)
- [ ] `margin-right` (need collapsing)
- [ ] `padding-left`
- [ ] `padding-right`
- [ ] `prefix`
- [ ] `postfix`
- [x] `text-decoration`

## `text` (inline)

## `blockquote` (block)

## `div` (block)

### Styles for div elements

- [ ] `after`
- [ ] `before`
- [x] `color`
- [ ] `fmt`
- [x] `font-style`
- [x] `font-weight`
- [ ] `margin-bottom` (need collapsing)
- [x] `margin-left`
- [x] `margin-right`
- [ ] `margin-top` (need collapsing)
- [x] `padding-bottom`
- [x] `padding-left`
- [x] `padding-right`
- [x] `padding-top`
- [x] `text-decoration`

## `dl` (block)

## `h1`, `h2`, `h3` (block)

## `li` (block)

## `meta` (block) (?)

## `ol` (block)

## `par` (block)

## `rule` (block)

## `status` (block) (?)

## `ul` (block)

## `verbatim` (block)

## `alert` (block)

## `alert_danger` (block)

## `alert_info` (block)

## `alert_success` (block)

## `alert_warning` (block)

## `bullets` (block)

# Component styling by style

## `after`

- [x] ~~text~~
- [x] sub
- [x] span
- [ ] div

## `background-color`

- [x] text
- [x] sub
- [x] span
- [x] div

## `before`

- [x] ~~text~~
- [x] sub
- [x] span
- [ ] div

## `class-map`

- [x] ~~text~~
- [ ] sub
- [x] ~~span~~
- [x] ~~div~~

## `color`

- [x] text
- [x] sub
- [x] span
- [x] div

## `collapse`

Called right after `transform`, it simplifies collapsing.

- [x] ~~text~~
- [x] sub
- [x] ~~span~~
- [x] ~~div~~

## `digits`

- [x] ~~text~~
- [ ] sub
- [x] ~~span~~
- [x] ~~div~~

## `fmt`

- [x] text
- [x] sub
- [x] span
- [x] div

## `font-style`

- [x] text
- [x] sub
- [x] span
- [x] div

## `font-weight`

- [x] text
- [x] sub
- [x] span
- [x] div

## `line-type`

- [x] ~~text~~
- [x] ~~sub~~
- [x] ~~span~~
- [x] ~~div~~

## `list-style-type`

- [x] ~~text~~
- [x] ~~sub~~
- [x] ~~span~~
- [x] ~~div~~

## `margin-bottom`

- [x] ~~text~~
- [x] ~~sub~~
- [x] ~~span~~
- [ ] div (need collapsing)

## `margin-left`

- [x] ~~text~~
- [x] ~~sub~~
- [ ] span (need collapsing)
- [x] div

## `margin-right`

- [x] ~~text~~
- [x] ~~sub~~
- [ ] span (need collapsing)
- [x] div

## `margin-top`

- [x] ~~text~~
- [ ] sub
- [ ] span
- [ ] div

## `padding-bottom`

- [x] ~~text~~
- [x] ~~sub~~
- [x] ~~span~~
- [x] div

## `padding-left`

- [x] ~~text~~
- [x] ~~sub~~
- [x] span
- [x] div

## `padding-right`

- [x] ~~text~~
- [x] ~~sub~~
- [x] span
- [x] div

## `padding-top`

- [x] ~~text~~
- [x] ~~sub~~
- [x] ~~span~~
- [x] div

## `prefix`

- [x] ~~text~~
- [x] sub
- [ ] span
- [ ] div

## `postfix`

- [x] ~~text~~
- [x] sub
- [ ] span
- [ ] div

## `start`

- [x] ~~text~~
- [x] ~~sub~~
- [x] ~~span~~
- [x] ~~div~~

## `string-quote`

- [x] ~~text~~
- [ ] sub
- [x] ~~span~~
- [x] ~~div~~

## `text-decoration`

- [x] text
- [x] sub
- [x] span
- [x] div

## `text-exdent`

- [x] ~~text~~
- [x] ~~sub~~
- [x] ~~span~~
- [ ] div

## `transform`

- [x] ~~text~~
- [x] sub
- [x] ~~span~~
- [x] ~~div~~

## `vec-last`

- [x] ~~text~~
- [x] sub
- [x] ~~span~~
- [x] ~~div~~

## `vec-sep`

- [x] ~~text~~
- [x] sub
- [x] ~~span~~
- [x] ~~div~~

## `vec-sep2`

- [x] ~~text~~
- [x] sub
- [x] ~~span~~
- [x] ~~div~~

## `vec-trunc`

- [x] ~~text~~
- [x] sub
- [x] ~~span~~
- [x] ~~div~~

## `vec-trunc-style`

- [x] ~~text~~
- [x] sub
- [x] ~~span~~
- [x] ~~div~~

# Styles cascading

## Automatic `span` elements and substitution

The tricky thing here is that some styles apply to the substitutions
inside the span, and not the span itself. In fact they might apply to
the elements of substituted vectors. (Not great, but this is what we need
to work with at this point.)

```{r}
fun <- function() {
  cli::cli_div(theme = list(span = list(after = "|", postfix = ">")))
  cli_text("foo {.emph bar} baz")
  cli_text("foo {.emph {1:4}} baz")
  cli_text("foo {.emph {1:4}, {5:8}} baz")
  cli_text("foo {.emph {1:4}xyz} baz")
  cli_text("foo {.emph {1:4}}{.emph xyz} baz")
}
fun()
```

Style inheritence also works differently if there is a single substitution
inside the span tag: then it is not applied to the tag itself.

So if we want to clear this up, then we need to keep the current styles,
and create new ones (possibly by qualifying the existing ones), that apply
either only to the span, or to the substitutions inside the span.

Possibly we need to define a new `expr` tag or qualifyier that we can use
in themes. With a tag we could have:
```
list("span expr" = list(after = ">"))
```
and with a new qualifier we could have
```
list(span = list("expr::after" = ">"))
```
The first one seems more logical, it is just a new tag, after all, no new
concept to learn, just that `{}` substitutions now create an implicit
`expr` tag.

We would need to keep the compatibility with the current themes, though, so
```
list(span = list(after = ">"))
```
would still apply to both the expressions and the tag itself, which is not
great. With qualifyers we could have
```
list(span = list("expr::after" = ">"))
list(span = list("self::after" = "**"))
```
or maybe
```
list(span = list("after::expr" = ">"))
list(span = list("after::self" = "**"))
```
where we explicitly choose whether to apply to the expression to the
substituted expressions (possibly elementwise), or to the tag itself.
