---
title: "Advanced cli progress bars"
author: "Gábor Csárdi"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    toc: true
    toc_depth: 2
editor_options:
  markdown:
    wrap: sentence
---

```{r, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  cache = TRUE
)
# Turn on ANSI colors
options(
  cli.num_colors = 256L
)
asciicast::init_knitr_engine(
  echo = TRUE,
  echo_input = FALSE,
  options = list(
    asciicast_at = NULL,
    asciicast_end_wait = 3,
    asciicast_speed = 1,
    asciicast_cpp11_linkingto = "[[cpp11::linking_to(\"cli\")]]"
  )
)
```

```{asciicast, setup, cache = FALSE, include = FALSE}
library(cli)
options(cli.progress_show_after = 0)
options(cli.progress_clear = FALSE)
options(cli.progress_format_iterator = NULL)
options(cli.progress_format_iterator_nototal = NULL)
options(cli.progress_format_tasks = NULL)
options(cli.progress_format_tasks_nototal = NULL)
options(cli.progress_format_download = NULL)
options(cli.progress_format_download_nototal = NULL)
```

```{asciicast}
library(cli)
```

# Overhead

cli progress bars do have an overhead that may or may not be significant
for your use case. In the R API, if you have a tight loop, then you should
not update the progress bar too often.

## Minimizing overhead

To minimize progress bar overhead, `cli_progress_update()` uses an internal
timer and only update the progress bar on the screen if the timer is due.

In C code, you can refer to the timer directly to avoid an update.
The `CLI_SHOULD_TICK` macro evaluates to one if the timer is due and an
update is needed, otherwise to zero. `CLI_SHOULD_TICK` only works if you
already created a cli progress bar from C, or you called
`cli_progress_init_timer()`. The latter initializes the cli timer without
creating a progress bar. (If the timer is not initialized, then
`CLI_SHOULD_TICK` evaluates to zero.)

```c
  SEXP bar = PROTECT(cli_progress_bar(num_iters, NULL));
  for (i = 0; i < num_iters; i++) {
    if (CLI_SHOULD_TICK) cli_progress_set(bar, i);
    // ...
  }
  cli_progress_done(bar);
```

# Non-interactive R sessions

Non-interactive R sessions typically have two differences to interactive
ones. First, by default the cli timer signals every
`r cli:::cli_timer_non_interactive` milliseconds, instead of
`r cli:::cli_timer_interactive` milliseconds.

Second, non-interactive sessions often do not have a dynamic terminal,
so cli cannot create and update a status bar in the last row of the
display. In this case cli simply prints every progress update as a new
line to the screen.

<!-- TODO: example -->

# Progress bars in scripts

You can use progress bars in R scripts, just like you use them in R
packages.

In an R script you might create a progress bar from the global environment,
instead from within a function call. The global environment also has a
_current_ progress bar, so when you create a progress bar from the global
environment, the previous one from the same environment is terminated.
However, there is not function to return from, so the last progress bar
of the script will be only terminated when the script terminates, or when
you explicitly terminate it using `cli_progress_done()`.

<!-- # Nested progress bars TODO -->

# Customization

cli progress bars can be customized by the developer and the end user,
by setting options, providing function arguments and regular cli themes.

Some aspects can only be customized by the developer, and some others can
only be customized by the end user. Others can be customized by both,
with the end user's setting taking precedence.

## Developer customization

### About progress bar types

Each progress bar type has a default display (format string), which can be
configured by the end user. The current progress bar types are, with their
default display, with known and unknown number of total progress units.

### `iterator`

Typically for loops and mapping functions. It shows a bar
by default, if the total number of iterations is known.

```{asciicast echo = FALSE,  R.options = list(asciicast_at = "end")}
cli_progress_demo("Data cleaning", total = 100, at = 50, clear = FALSE)
```

```{asciicast echo = FALSE, R.options = list(asciicast_at = "end")}
cli_progress_demo("Data cleaning", at = 50, clear = FALSE)
```

### `tasks`

For a list of tasks, by default it shows a `current/total` display.

```{asciicast echo = FALSE, R.options = list(asciicast_at = "end")}
cli_progress_demo(
  "Finding data files", total = 100, at = 50,
  clear = FALSE, type = "tasks"
)
```

```{asciicast echo = FALSE, R.options = list(asciicast_at = "end")}
cli_progress_demo(
  "Finding data files", at = 50,
  clear = FALSE, type = "tasks"
)
```

### `download`

For downloads, progress units are shown as bytes by default here.

```{asciicast echo = FALSE, R.options = list(asciicast_at = "end")}
cli_progress_demo(
  "Downloading", total = 10280, at = 5120, clear = FALSE,
  type = "download"
)
```

```{asciicast echo = FALSE, R.options = list(asciicast_at = "end")}
cli_progress_demo(
  "Downloading", at = 5120, clear = FALSE,
  type = "download"
)
```

### `custom`

For custom displays, the developer has to specify an format string
for `custom` progress bars.

### Custom format strings (by the developer)

The developer can specify a custom format string for a progress bar.
For `custom` progress bars, this is compulsory. Format strings may use glue
templating, cli pluralization and cli theming. They can also use a number
of built-in cli progress variables, see 'Progress variables' below.

```{asciicast}
f <- function() {
  cli_progress_bar(
    total = 20000,
    format = "Step {step} | {pb_bar} {pb_percent}"
  )
  step <- 1
  for (i in 1:10000) {
    Sys.sleep(2/10000)
    cli_progress_update(set = i)
  }
  step <- 2
  for (i in 10001:20000) {
    Sys.sleep(2/10000)
    cli_progress_update(set = i)
  }
}
f()
```

For `custom` progress bars cli always uses the specified format string.
For other types, the end user might customize the format string, see below.

## End user customization

### Quick loops

The `cli.progress_show_after` (default is two seconds) option is the number
seconds to wait before showing a progress bar. 

### Custom bars

The end user can customize how a progress bar will look, by setting one
or more of the following options:

* `cli.progress_bar_style`
* `cli.progress_bar_style_unicode`
* `cli.progress_bar_style_ascii`

On UTF-8 displays `cli.progress_bar_style_unicode` is used, if set.
Otherwise `cli.progress_bar_style` is used. On non UTF-8 displays
`cli.progress_bar_style_ascii` is used, if set. Otherwise
`cli.progress_bar_style` is used.

These options can be set to a built-in progress bar style name:

```{r include = FALSE}
library(cli)
```

```{r}
names(cli_progress_styles())
```

```{asciicast}
options(cli.progress_bar_style = "fillsquares")
f <- function() lapply(cli_progress_along(letters), function(l) Sys.sleep(0.2))
x <- f()
```

```{asciicast include = FALSE, cache = FALSE}
options(cli.progress_bar_style = NULL)
```

Alternatively, they can be set to a list with entries `complete`,
`incomplete` and `current`, to specify the characters (or strings) for the
parts of the progress bar:

```{asciicast}
options(cli.progress_bar_style = list(
  complete = cli::col_yellow("\u2605"),
  incomplete = cli::col_grey("\u00b7")
))
f <- function() lapply(cli_progress_along(letters), function(l) Sys.sleep(0.2))
x <- f()
```

```{asciicast include = FALSE}
options(cli.progress_bar_style = NULL)
```

### Custom spinners

Options to customize cli spinners:

* `cli.spinner`
* `cli.spinner_unicode`
* `cli.spinner_ascii`

On UTF-8 displays `cli.spinner_unicode` is used, if set, otherwise
`cli.spinner`. On ASCII displays `cli.spinner_ascii` is used, if set,
otherwise `cli.spinner`.

Use `list_spinners()` to list all spinners and `demo_spinners()` to take
a peek at them.

```{asciicast custom-spinner}
options(cli.spinner = "moon")
f <- function() {
  cli_progress_bar(format = strrep("{cli::pb_spin} ", 20), clear = TRUE)
  for (i in 1:100) {
    Sys.sleep(5/100)
    cli_progress_update()
  }
}
f()
```

### Custom format strings

The end user may use a number of global options to customize how the
built-in progress bar types are displayed on the screen:

* `cli.progress_format_iterator` is used for `iterator` progress bars.
* `cli.progress_format_iterator_nototal` is used for `iterator` progress
   bars with an unknown number of total units.
* `cli.progress_format_tasks` is used for `tasks` progress bars.
* `cli.progress_format_tasks_nototal` is used for `tasks` progress
   bars with an unknown number of total units.
* `cli.progress_format_download` is used for `download` progress bars.
* `cli.progress_format_download_nototal` is used for `download` progress
   bars with an unknown number of total units.


## Progress variables

Custom format strings may use progress variables in glue interpolated
expressions, to refer to the state of the progress bar. See
`?"progress-variables"` in the manual for the list of supported variables.

If you refer to a progress variable from a package, you need need to
import it or qualify the reference with `cli::`. When you set a custom
format string as an end user option, we suggest that you always use the
qualified form, in case the cli package is not attached. For example, to
set a minimal display for downloads you might write 

```{asciicast include = FALSE, cache = FALSE}
options(cli.progress_format_download =
  paste0(
    "{cli::col_cyan('\u2B07')} {cli::pb_spin} ",
    "{cli::pb_name}[{cli::pb_current_bytes}/{cli::pb_total_bytes}]"
  )
)
```

```{asciicast echo = FALSE, R.options = list(asciicast_at = "end")}
cli_progress_demo(
"Downloading", total = 10280, at = 5121, clear = FALSE,
  type = "download"
)
```

You can use your own expressions and functions on progress bar tokens.
E.g. to show the current number of steps with letters instead of numbers,
use `letters[pb_current]`:

```{asciicast function-of-token}
f <- function() {
  cli_progress_bar(
    total = 26,
    format = "{pb_spin} This is step {.emph {letters[pb_current]}}. {pb_spin}"
  )
  for (i in 1:26) {
    Sys.sleep(3/26)
    cli_progress_update()
  }
}
f()
```

## Clearing or keeping terminated progress bars

By default terminated progress bars are removed from the screen. The end
user can set the `cli.progress_clear` option to `FALSE` to override the
default. In addition, the developer can also change the default, using the
`clear` parameter of `cli_progress_bar()`. If both the option and the
parameter are set, the parameter is used.

# The C API

To use the C cli progress
API in your package, you need to add cli to `LinkingTo` and `Imports`:
```
LinkingTo: cli
Imports: cli
```

In the C files you want to use the API from include `cli/progress.h`:

```c
#include <cli/progress.h>
```

Now you are ready to call cli functions. The C API is similar to the
traditional R API:

1. `cli_progress_bar()` creates a progress bar.
2. `cli_progress_update()` updates a progress bar.
3. `cli_progress_done()` terminates it.

A complete example:

```{asciicastcpp11 capi}
#include <cli/progress.h>
SEXP progress_test1() {
  int i;
  SEXP bar = PROTECT(cli_progress_bar(1000, NULL));
  for (i = 0; i < 1000; i++) {
    cli_progress_sleep(0, 4 * 1000 * 1000);
    if (CLI_SHOULD_TICK) cli_progress_set(bar, i);
  }
  cli_progress_done(bar);
  UNPROTECT(1);
  return Rf_ScalarInteger(i);
}
```

```{asciicast, echo = FALSE, dependson = -1}
invisible(progress_test1())
```

## C API reference

```{r include = FALSE, cache = FALSE, child = cli:::docs_progress_c_api()}
```
