---
title: "Building a Semantic CLI"
author: "Gábor Csárdi"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    toc: true
    toc_depth: 2
editor_options: 
  markdown: 
    wrap: sentence
---

```{r, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  cache = TRUE
)
# Turn on ANSI colors
options(cli.num_colors = 256L)
asciicast::init_knitr_engine(
  startup = quote({
    library(cli)
    set.seed(1) }),
  echo = TRUE,
  echo_input = FALSE)
```

# Introduction

Starting from version 2.0.0, the cli package can build a semantic Command Line Interface (CLI).
By *semantic* we mean that `cli_*` functions that create console output are focused on the role of the text in the CLI, as opposed to how exactly it is formatted.
The formatting of the CLI can be specified separately, and can be also customized by the end user.
This is not unlike how HTML and CSS works as well, so it might be already familiar.

The semantic `cli_*` functions create semantic elements: headings, lists, alerts, paragraphs, etc.
In this introduction we will go over these functions and also some common features of them.

# Building a command line interface

```{r, setup}
library(cli)
```

To build a CLI, you can simply start using the `cli_*` functions to create various CLI elements.
Their exact formatting depends on the current theme, see 'Theming' below.

## Alerts

Alerts are typically short messages.
cli has four types of alerts (success, info, warning, danger) and also a generic alert type:

```{asciicast}
cli_alert_success("Updated database.")
```

```{asciicast}
cli_alert_info("Reopened database.")
```

```{asciicast}
cli_alert_warning("Cannot reach GitHub, using local database cache.")
```

```{asciicast}
cli_alert_danger("Failed to connect to database.")
```

```{asciicast}
cli_alert("A generic alert")
```

## Text

Text is automatically wrapped to the terminal width.

```{asciicast}
cli_text(cli:::lorem_ipsum())
```

## Paragraphs

Paragraphs break the output.
The default theme inserts an empty line before and after paragraphs, but only a single empty line is added between two paragraphs.

```{asciicast}
fun <- function() {
  cli_par()
  cli_text("This is some text.")
  cli_text("Some more text.")
  cli_end()
  cli_par()
  cli_text("Already a new paragraph.")
  cli_end()
}
fun()
```

`cli_end()` closes the latest open paragraph (or other open container).

## Auto-closing containers

If a paragraph (or other container, see 'Generic containers' later), is opened within a function, cli automatically closes it at the end of the function, by default.
So in the previous example the last `cli_end()` call is not needed.
Use `.auto_close = FALSE` in `cli_par()` to leave the paragraph open after its calling function returns.

## Headings

cli suppports three levels of headings.
This is how they look in the default theme.
The default theme adds an empty line before headings, and an empty line after `cli_h1()` and `cli_h2()`.

```{asciicast}
cli_h1("Heading 1")
```

```{asciicast}
cli_h2("Heading 2")
```

```{asciicast}
cli_h3("Heading 3")
```

## Interpolation

All cli text is treated as a glue template, with special formatters available (see the 'Inline text formatting' Section):

```{asciicast}
size <- 123143123
dt <- 1.3454
cli_alert_info(c(
  "Downloaded {prettyunits::pretty_bytes(size)} in ",
  "{prettyunits::pretty_sec(dt)}"))
```

## Inline text formatting

To define inline markup, you can use the regular glue braces, and after the opening brace, supply the name of the markup formatter with a leading dot, e.g. for emphasised text, you use `.emph`.
Some examples are below, see `?"inline-markup"` for details.

```{asciicast}
fun <- function() {
  cli_ul()
  cli_li("{.emph Emphasized} text")
  cli_li("{.strong Strong} importance")
  cli_li("A piece of code: {.code sum(a) / length(a)}")
  cli_li("A package name: {.pkg cli}")
  cli_li("A function name: {.fn cli_text}")
  cli_li("A keyboard key: press {.kbd ENTER}")
  cli_li("A file name: {.file /usr/bin/env}")
  cli_li("An email address: {.email bugs.bunny@acme.com}")
  cli_li("A URL: {.url https://acme.com}")
  cli_li("An environment variable: {.envvar R_LIBS}")
  cli_li("Some {.field field}")
}
fun()
```

To combine inline markup and string interpolation, you need to add another set of braces:

```{asciicast}
dlurl <- "https://httpbin.org/status/404"
cli_alert_danger("Failed to download {.url {dlurl}}.")
```

`"val"` is a special inline style, that in the default theme calls `cli_format()` to tailor the conversion of values to strings.
The conversion can be themed, see "Theming" below.

```{asciicast}
cli_div(theme = list(.val = list(digits = 2)))
cli_text("Some random numbers: {.val {runif(4)}}.")
cli_end()
```

## Inline lists of items

When cli performs inline text formatting, it automatically collapses glue substitutions, after formatting.
This is handy to create lists of files, packages, etc.

```{asciicast}
pkgs <- c("pkg1", "pkg2", "pkg3")
cli_text("Packages: {pkgs}.")
```

```{asciicast}
pkgs <- c("pkg1", "pkg2", "pkg3")
cli_text("Packages: {.pkg {pkgs}}")
```

By default class names are collapsed differently:

```{asciicast}
x <- Sys.time()
cli_text("Hey {.var x} has class {.cls {class(x)}}")
```

## Non-breaking spaces

Use `\u00a0` to create a non-breaking space.
E.g. in here we insert some non-breaking spaces, and mark them with an `X`, so it is easy to see that there are no line breaks at a non-breaking space:

```{asciicast}
# Make some spaces non-breaking, and mark them with X
txt <- cli:::lorem_ipsum()
mch <- gregexpr(txt, pattern = " ", fixed = TRUE)
nbs <- runif(length(mch[[1]])) < 0.5
regmatches(txt, mch)[[1]] <- ifelse(nbs, "X\u00a0", " ")
cli_text(txt)
```

## Lists

cli has three types of list: ordered, unordered and definition lists, see `cli_ol()`, `cli_ul()` amd `cli_dl()`:

```{asciicast}
cli_ol(c("item 1", "item 2", "item 3"))
```

```{asciicast}
cli_ul(c("item 1", "item 2", "item 3"))
```

```{asciicast}
cli_dl(c("item 1" = "description 1", "item 2" = "description 2",
         "item 3" = "description 3"))
```

Item text is wrapped to the terminal width:

```{asciicast}
cli_ul(c("item 1" = cli:::lorem_paragraph(1, 20),
         "item 2" = cli:::lorem_paragraph(1, 20)))
```

### Adding list items iteratively

Items can be added one by one:

```{asciicast}
fun <- function() {
  lid <- cli_ul()
  cli_li("Item 1")
  cli_li("Item 2")
  cli_li("Item 3")
  cli_end(lid)
}
fun()
```

The `cli_ul()` call creates a list container, and because its items are not specified, it leaves the container open.
Then items can be added one by one.
(The last `cli_end()` is not necessary, because by default containers auto-close when their calling function exits.)

### Adding text to an item iteratively

`cli_li()` creates a new container for the list item, within the list container.
You can keep adding text to the item, until the container is closed via `cli_end()` or a new `cli_li()`, which closes the current item container, and creates another one for the new item:

```{asciicast}
fun <- function() {
  cli_ul()
  cli_li("First item")
  cli_text("This is still the first item")
  cli_li("This is the second item")
}
fun()
```

### Nested lists

To create nested lists, open nested containers:

```{asciicast}
fun <- function() {
  cli_ol()
  cli_li("Item 1")
  ulid <- cli_ul()
  cli_li("Subitem 1")
  cli_li("Subitem 2")
  cli_end(ulid)
  cli_li("Item 2")
  cli_end()
}
fun()
```

In `cli_end(olid)`, the `olid` is necessary, otherwise `cli_end()` would only close the container of the list item.

## Horizontal rules

`cli_rule()` creates a horizontal rule.

```{asciicast}
cli_rule(left = "Compiling {.pkg mypackage}")
```

You can use the usual inline markup in the labels of the rule.
The rule's appearance is specified in the current theme.
In particular:

-   `before` is added before the rule.
-   `after` is added after the rule.
-   `color` is used for the color of the rule *and* the labels. (Use color within the label text for a different label color.)
-   `background-color`: is the background color for the rule and the labels. (Again, you can use a different background color within the label itself.)
-   `margin-top`, `margin-bottom` for empty space above and below the rule.
-   `line-type` specifies the line type of the rule. See `?cli_rule` for line types.

## The status bar

cli supports creating a status bar in the last line of the console, if the terminal supports the carriage return control character to move the cursor to the beginning of the line.
This is supported in all terminals, in RStudio, Emacs, RGui, R.app, etc.
It is not supported if the output is a file, e.g. typically on CI systems.

`cli_process_start()` creates a new status and shows it in the status bar and `cli_process_update()` updates a status.
By default cli automatically terminates the status and updates the status bar accordingly.

While it is possible to create multiple status bars, on a typical terminal only one of them can be shown at any time.
To reduce the number of calls to cli and thus clutter in the code, cli by default terminates a status if another one is created from the same caller.

While the status bar is active, cli can still produce output, as normal.
This output is created "above" the status bar, which is always kept in the last line of the screen.
See the following example:

```{asciicast, R.options = list(asciicast_at = NULL, asciicast_end_wait = 30, asciicast_rows = 3)}
f <- function(n = 10) {
  cli_alert_info("About to start downloads of {n} file{?s}")
  i <- 0
  sb <- cli_process_start("Got {i}/{n} {qty(i)}file{?s}.")
  for (i in seq_len(n)) {
    Sys.sleep(0.5)
    if (i == 5) cli_alert_info("Already half way!")
    cli_process_update()
  }
}
f()
```

# Theming

The looks of the various CLI elements can be changed via themes.
The cli package comes with a simple built-in theme, and new themes can be added as well.

## Tags, ids and classes

Similarly to HTML document, the elements of a CLI form a tree of nodes.
Each node has exactly one tag, at most one id, and optionally a set of classes.
E.g. `cli_par()` creates a node with a `<p>` tag, `cli_ol()` creates a node with an `<ol>` tag, etc.
Here is an example CLI tree.
It always starts with a `<body>` tag with id `"body"`, this is created automatically.

    <body id="body">
      <par>
        <ol>
          <it>
            <span class="pkg">

A cli theme is a named list, where the names are selectors based on tag names, ids and classes, and the elements of the list are style declarations.
For example, the style of `<h1>` tags looks like this in the built-in theme:

```{asciicast}
builtin_theme()$h1
```

See also `?cli::themes` for the reference and `?cli::simple_theme` for an example theme.

## Generic containers

`cli_div()` is a generic container, that does not produce any output, but it can add a new theme.
This theme is removed when the `<div>` node is closed.
(Like other containers, `cli_div()` auto-closes when the calling function exits.)

```{asciicast}
fun <- function() {
  cli_div(theme = list (.alert = list(color = "red")))
  cli_alert("This will be red")
  cli_end()
  cli_alert("Back to normal color")
}
fun()
```

## Theming inline markup

The inline markup formatters always use a `<span>` tag, and add the name of the formatter as a class.

```{asciicast}
fun <- function() {
  cli_div(theme = list(span.emph = list(color = "orange")))
  cli_text("This is very {.emph important}")
  cli_end()
  cli_text("Back to the {.emph previous theme}")
}
fun()
```

In addition to adding inline markup explicitly, like `.emph` here, cli can use the class(es) of the substituted expression to style it automatically.
This can be configured as part of the theme, in the form of a mapping from the `class()` of the expression, to the name of the markup formatter.
For example, if we have a `filename` S3 class, we can make sure that it is always shown as a `.file` in the cli output:

```{asciicast}
fun <- function() {
  cli_div(theme = list(body = list("class-map" = list("filename" = "file"))))
  fns <- structure(c("file1", "file2", "file3"), class = "filename")
  cli_text("Found some files: {fns}.")
  cli_end()
}
fun()
```

# CLI messages

All `cli_*()` functions are implemented using standard R conditions.
For example a `cli_alert()` call emits an R condition with class `cli_message`.
These messages can be caught, muffled, transfered from a subprocess to the main R process.

When a cli function is called:

1.  cli throws a `cli_message` condition.
2.  If this condition is caught and muffled (via the `cli_message_handled` restart), then nothing else happens.
3.  Otherwise the `cli.default_handler` option is checked and if this is a function, then it is called with the message.
4.  If the `cli.default_handler` option is not set, or it is not a function, the default cli handler is called, which shows the text, alert, heading, etc. on the screen, using the standard R `message()` function.

```{asciicast}
tryCatch(cli_h1("Heading"), cli_message = function(x) x)
suppressMessages(cli_text("Not shown"))
```

# Subprocesses

If `cli_*()` commands are invoked in a subprocess via `callr::r_session` (see <https://callr.r-lib.org>), and the `cli.message_class` option is set to `"callr_message"`, then cli messages are automatically copied to the main R process:

```{asciicast}
rs <- callr::r_session$new()
rs$run(function() {
  options(cli.message_class = "callr_message")
  cli::cli_text("This is subprocess {.emph {Sys.getpid()}} from {.pkg callr}")
  Sys.getpid()
})
invisible(rs$close())
```

# Utility functions

## ANSI colors

cli has functions to create ANSI colored or styled output in the console.
`col_*` functions change the foreground color, `bg_*` functions change the background color, and `style_*` functions change the style of the text in some way.

These functions concatenate their arguments using `paste0()`, and add the `ansi_string` class to their result:

```{asciicast}
cat(col_red("This ", "is ", "red."), sep = "\n")
```

Foreground colors:

```{asciicast}
cli_ul(c(
  col_black("black"),
  col_blue("blue"),
  col_cyan("cyan"),
  col_green("green"),
  col_magenta("magenta"),
  col_red("red"),
  col_white("white"),
  col_yellow("yellow"),
  col_grey("grey")
))
```

Note that these might actually look different depending on your terminal theme.
Background colors:

```{asciicast}
cli_ul(c(
  bg_black("black background"),
  bg_blue("blue background"),
  bg_cyan("cyan background"),
  bg_green("green background"),
  bg_magenta("magenta background"),
  bg_red("red background"),
  bg_white("white background"),
  bg_yellow("yellow background")
))
```

Text styles:

```{asciicast}
cli_ul(c(
  style_dim("dim style"),
  style_blurred("blurred style"),
  style_bold("bold style"),
  style_hidden("hidden style"),
  style_inverse("inverse style"),
  style_italic("italic style"),
  style_reset("reset style"),
  style_strikethrough("strikethrough style"),
  style_underline("underline style")
))
```

Not all `style_*` functions are supported by all terminals.

Colors, background colors and styles can be combined:

```{asciicast}
bg_white(style_bold(col_red("TITLE")))
```

`make_ansi_style()` can create custom colors, assuming your terminal supports them.
`combine_ansi_styles()` combines several styles into a function:

```{asciicast}
col_warn <- combine_ansi_styles(make_ansi_style("pink"), style_bold)
col_warn("This is a warning in pink!")
cat(col_warn("This is a warning in pink!"))
```

## Console capabilities

Query the console width:

```{asciicast}
console_width()
```

Query if the console supports ansi escapes:

```{asciicast}
is_ansi_tty()
```

Hide the cursor, if the console supports it (no-op otherwise):

```{r}
ansi_hide_cursor()
ansi_show_cursor()
```

See also `ansi_with_hidden_cursor()`.

Query if the console supports `\r`:

```{asciicast}
is_dynamic_tty()
```

Query if the console supports UTF-8 output:

```{asciicast}
is_utf8_output()
```

## Unicode characters

The `symbol` variable includes some Unicode characters that are often useful in CLI messages.
They automatically fall back to ASCII symbols if the platform does not support them.
You can use these symbols both with the semantic `cli_*()` functions and directly.

```{asciicast}
cli_text("{symbol$tick} no errors  |  {symbol$cross} 2 warnings")
```

Here is a list of all symbols:

```{asciicast}
list_symbols()
```

Most symbols were inspired by (and copied from) the awesome [figures](https://github.com/sindresorhus/figures) JavaScript project.

## Spinners

See `list_spinners()` and `get_spinner()`.
From the awesome [cli-spinners](https://github.com/sindresorhus/cli-spinners#readme) project.

```{asciicast}
list_spinners()
```

```{asciicast}
get_spinner("dots")
```

```{asciicast, R.options = list(asciicast_at = NULL)}
ansi_with_hidden_cursor(demo_spinners("dots"))
```

```{asciicast, R.options = list(asciicast_at = NULL)}
ansi_with_hidden_cursor(demo_spinners("clock"))
```
