---
title: "From usethis::ui functions to cli"
author: "Gábor Csárdi"
date: "`r Sys.Date()`"
description: >
  Transition from usethis to cli.
output:
  rmarkdown::html_document:
    toc: true
    toc_depth: 2
editor_options: 
  markdown: 
    wrap: sentence
---

```{r, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  cache = TRUE
)
asciicast::init_knitr_engine(
  startup = quote({
    library(cli)
    library(usethis)
    set.seed(1)
  })
)
```

```{r setup}
library(cli)
library(usethis)
```

# Introduction

We'll show how to transition from the `usethis::ui_*` functions to cli 2.0.0.

# How to

## `usethis::ui_code()`

    usethis::ui_code(x)

```{asciicast}
ui_todo("Redocument with {ui_code('devtools::document()')}")
```

In general inline code formatting can be done with inline styles in cli.
The default theme has a `"code"` class, but it also one for functions, so this can be either of:

```{asciicast}
cli_ul("Redocument with {.code devtools::document()}")
```

```{asciicast}
cli_ul("Redocument with {.fun devtools::document}")
```

## `usethis::ui_code_block()`

    usethis::ui_code_block(x, copy = interactive(), .envir = parent.frame())

```{asciicast}
ui_code_block("{format(cli_code)}")
```

`cli_code()` produces similar output and it also syntax highlight R code:

```{asciicast}
cli_code(format(cli_code))
```

However, cli does not copy stuff to the clipboard, so this has to be done separately.

Another difference is that it also does not run glue substitutions on the code text, so if you want that to happen you'll need to do that before the cli call.

## `usethis::ui_done()`

    usethis::ui_done(x, .envir = parent.frame())

```{asciicast}
name <- "VignetteBuilder"
value <- "knitr, rmarkdown"
ui_done("Setting {ui_field(name)} field in DESCRIPTION to {ui_value(value)}")
```

This is probably closest to `cli_alert_success()`:

```{asciicast}
cli_alert_success("Setting {.field {name}} field in DESCRIPTION to {.val {value}}")
```

If you want to handle success and failure, then maybe the `cli_process_*()` functions are a better fit:

```{asciicast}
#| asciicast_at = "all",
#| asciicast_end_wait = 5
tryCatch({
    cli_process_start("Setting {.field {name}} field in DESCRIPTION to {.val {value}}")
    Sys.sleep(1) # <- do the task here, we just sleep
    cli_process_done() },
  error = function(err) {
    cli_process_failed()
    cli_alert_danger("Failed to ...")
  }
)
```

## `usethis::ui_field()`

    usethis::ui_field(x)

```{asciicast}
name <- "VignetteBuilder"
value <- "knitr, rmarkdown"
ui_done("Setting {ui_field(name)} field in DESCRIPTION to {ui_value(value)}")
```

cli has a `"field"` class for inline styling:

```{asciicast}
cli_alert_success("Setting {.field {name}} field in DESCRIPTION to {.val {value}}")
```

Just like `usethis::ui_field()` and similar usethis functions, cli collapses inline vectors, before styling:

```{asciicast}
name <- c("Depends", "Imports", "Suggests")
ui_done("Setting the {ui_field(name)} field(s) in DESCRIPTION")
```

```{asciicast}
cli_alert_success("Setting the {.field {name}} field{?s} in DESCRIPTION")
```

cli also helps you with the correct pluralization:

```{asciicast}
name <- "Depends"
cli_alert_success("Setting the {.field {name}} field{?s} in DESCRIPTION")
```

## `usethis::ui_info()`

    usethis::ui_info((x, .envir = parent.frame())

```{asciicast}
ui_info("No labels need renaming")
```

This is simply `cli_alert_info()`:

```{asciicast}
cli_alert_info("No labels need renaming")
```

## `usethis::ui_line()`

    usethis::ui_line(x, .envir = parent.frame())

```{asciicast}
ui_line("No matching issues/PRs found.")
```

This is just a line of text, so `cli_text()` is fine for this.
One difference is that `cli_text()` will automatically wrap the long lines.

```{asciicast}
cli_text("No matching issues/PRs found.")
```

## `usethis::ui_nope()`

    ui_nope(x, .envir = parent.frame())

cli does not support user input currently, so this has to stay in usethis.

## `usethis::ui_oops()`

    usethis::ui_oops(x, .envir = parent.frame())

```{asciicast}
ui_oops("Can't validate token. Is the network reachable?")
```

This is mostly just `cli_alert_danger()`, but for see also the `cli_process_*()` alternatives at `usethis::ui_done()`.

## `usethis::ui_path()`

    usethis::ui_path(x, base = NULL)

`ui_path()` formats paths as relative to the project or the supplied base directory, and also appends a `/` to directories.
cli does not do any of these, but it does have inline markup for files and paths:

```{asciicast}
logo_path <- file.path("man", "figures", "logo.svg")
img <- "/tmp/some-image.svg"
ui_done("Copied {ui_path(img)} to {ui_path(logo_path)}")
```

```{asciicast}
cli_alert_success("Copied {.file {img}} to {.file {logo_path}}")
```

## `usethis::ui_stop()`

    usethis::ui_stop(x, .envir = parent.frame())

`ui_stop()` does glue substitution on the string, and then calls `stop()` to throw an error.
`cli_abort()` does the same and is formatted using `cli_bullets()`.

```{asciicast}
ui_stop("Could not copy {ui_path(img)} to  {ui_path(logo_path)}, file already exists")
```

```{asciicast}
cli_abort(
  "Could not copy {.file {img}} to {.file {logo_path}}, file already exists",
  "i" = "You can set {.arg overwrite = TRUE} to avoid this error"
  )
```

## `usethis::ui_todo()`

    usethis::ui_todo(x, .envir = parent.frame())

This is a bullet, so either `cli_ul()` or `cli_alert_info()` should be appropriate:

```{asciicast}
ui_todo("Redocument with {ui_code('devtools::document()')}")
```

```{asciicast}
cli_ul("Redocument with {.fun devtools::document}")
```

## `usethis::ui_value()`

    usethis::ui_value(x)

The `"value"` inline class is appropriate for this.

```{asciicast}
name <- "VignetteBuilder"
value <- "knitr, rmarkdown"
ui_done("Setting {ui_field(name)} field in DESCRIPTION to {ui_value(value)}")
```

```{asciicast}
cli_alert_success("Setting {.field {name}} field in DESCRIPTION to {.val {value}}")
```

## `usethis::ui_warn()`

    usethis::ui_warn(x, .envir = parent.frame())

`ui_warn()` does glue substitution on the string, and then calls `warning()` to throw a warning.
`cli_war()` does the same and is formatted using `cli_bullets()`.

```{asciicast}
ui_warn("Could not copy {ui_path(img)} to  {ui_path(logo_path)}, file already exists")
```

```{asciicast}
cli_warn(
  "Could not copy {.file {img}} to {.file {logo_path}}, file already exists",
  "i" = "You can set {.arg overwrite = TRUE} to avoid this warning"
  )
```

## `usethis::ui_yeah()`

    ui_yeah(x, .envir = parent.frame())

cli does not support user input currently, so this has to stay in usethis.
