
```{r, setup, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  tidy = FALSE,
  error = FALSE,
  fig.path = "inst/figure/")
```

# boxes

> Draw Boxes in the R Console

[![Linux Build Status](https://travis-ci.org/r-lib/boxes.svg?branch=master)](https://travis-ci.org/r-lib/boxes)
[![Windows Build status](https://ci.appveyor.com/api/projects/status/github/r-lib/boxes?svg=true)](https://ci.appveyor.com/project/gaborcsardi/boxes)
[![](http://www.r-pkg.org/badges/version/boxes)](http://www.r-pkg.org/pkg/boxes)
[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/boxes)](http://www.r-pkg.org/pkg/boxes)
[![Coverage Status](https://img.shields.io/codecov/c/github/r-lib/boxes/master.svg)](https://codecov.io/github/r-lib/boxes?branch=master)

It integrates well with `crayon` for coloring the boxes or their content,
and `clisymbols` to include Unicode characters in the boxes. It can also
draw rules to make headers or footers in console output.

## Installation

```r
devtools::install_github("r-lib/boxes")
```

## Usage

```{r}
library(boxes)
```

```{r echo = FALSE, results = "hide"}
options(crayon.enabled = TRUE)
options(crayon.colors = 256)
crayon::num_colors(forget = TRUE)
```

### Rules

Simple rule

```{r rule, render=ansistrings:::knit_print.html}
rule()
```

Double rule

```{r rule-double, render=ansistrings:::knit_print.html}
rule(line = 2)
```

Bars

```{r rule-bars, render=ansistrings:::knit_print.html}
rule(line = "bar2")
rule(line = "bar5")
```

Left label

```{r rule-left, render=ansistrings:::knit_print.html}
rule(left = "Results")
```

Centered label

```{r rule-center, render=ansistrings:::knit_print.html}
rule(center = " * RESULTS * ")
```

Colored labels

```{r rule-color, render=ansistrings:::knit_print.html}
rule(center = crayon::red(" * RESULTS * "))
```

Colored line

```{r rule-color-line, render=ansistrings:::knit_print.html}
rule(center = crayon::red(" * RESULTS * "), line_color = "red")
```

Custom line

```{r rule-line-custom, render=ansistrings:::knit_print.html}
rule(center = "TITLE", line = "~")
```

More custom line

```{r rule-line-custom-2, render=ansistrings:::knit_print.html}
rule(center = "TITLE", line = crayon::blue("~-"))
```
Even more custom line

```{r rule-line-custom-3, render=ansistrings:::knit_print.html}
rule(center = crayon::bgRed(" ", clisymbols::symbol$star, "TITLE",
  clisymbols::symbol$star, " "),
  line = "\u2582",
  line_color = "orange")
```

### Boxes

Default box

```{r box, render=ansistrings:::knit_print.html}
boxx("Hello there!")
```

Change border style

```{r box-border-style, render=ansistrings:::knit_print.html}
boxx("Hello there!", border_style = "double")
```

Multiple lines

```{r box-lines, render=ansistrings:::knit_print.html}
boxx(c("Hello", "there!"), padding = 1)
```

Padding

```{r box-padding, render=ansistrings:::knit_print.html}
boxx("Hello there!", padding = 1)
boxx("Hello there!", padding = c(1, 5, 1, 5))
```

Margin

```{r box-margin, render=ansistrings:::knit_print.html}
boxx("Hello there!", margin = 1)
boxx("Hello there!", margin = c(1, 5, 1, 5))
boxx("Hello there!", padding = 1, margin = c(1, 5, 1, 5))
```

Floating

```{r box-floating, render=ansistrings:::knit_print.html}
boxx("Hello there!", padding = 1, float = "center")
boxx("Hello there!", padding = 1, float = "right")
```

Text color

```{r box-color, render=ansistrings:::knit_print.html}
boxx(crayon::cyan("Hello there!"), padding = 1, float = "center")
```

Backgorund color

```{r box-bgcolor, render=ansistrings:::knit_print.html}
boxx("Hello there!", padding = 1, background_color = "brown")
boxx("Hello there!", padding = 1, background_color = crayon::bgRed)
```

Border color

```{r box-border-color, render=ansistrings:::knit_print.html}
boxx("Hello there!", padding = 1, border_color = "green")
boxx("Hello there!", padding = 1, border_color = crayon::red)
```

Label alignment

```{r box-label-align, render=ansistrings:::knit_print.html}
boxx(c("Hi", "there", "you!"), padding = 1, align = "left")
boxx(c("Hi", "there", "you!"), padding = 1, align = "center")
boxx(c("Hi", "there", "you!"), padding = 1, align = "right")
```

A very customized box

```{r box-customized, render=ansistrings:::knit_print.html}
star <- clisymbols::symbol$star
label <- c(paste(star, "Hello", star), "  there!")
boxx(
  crayon::white(label),
  border_style="round",
  padding = 1,
  float = "center",
  border_color = "tomato3",
  background_color="darkolivegreen"
)
```

## License

MIT © RStudio
